% !TEX root = ../../thesis.tex
\newpage
%\pagestyle{thesis}
\acknowledgements

\vspace{-0.25\baselineskip}
This work is the culmination of a lifelong arc that would not have been possible without the support of many people, all of whom deserve many thanks. First, those who guided me along my mathematical journey: my high school teachers Donna Fallon and Karen Petramale for their tremendous enthusiasm, my college professors Dr. David Brown, Dr. James Conklin, Dr. Michael `Bodhi' Rogers, Dr. Teresa Moore, Dr. Matthew Price, Dr. Emilie Wiesner for their endless time and consideration, especially Professor Victor Symonette without whom I may not be here today. Finally, the faculty at Syracuse University: Dr. Uday Banerjee, Dr. Gerald Cargo, Dr. Steven Diaz, Dr. Duane Graysay, Dr. Thomas John, Dr. Leonid Kovalev, Dr. Loredana Lanzani, Dr. Graham Leuschke, Dr. Moira McDermott, Dr. Jeffrey Meyer, Dr. Claudia Miller, Dr. Dan Zacharia in addition to Kim Canino, Kelly Jarvi, Julie O'Connor, Sandra Ware for all their time and guidance. I would also like to thank Professor John Voight and Professor David Zywina for our short but helpful conversations. 

\vspace{-0.25\baselineskip}
Of course, I would not have had the perseverance to have made it through a Ph.D. program without all the happiness that my colleagues and friends, especially at Syracuse University, brought to the process: Andrew and Morgan Ascanio, Will Carrara, Paul Casler, Ben Dows, Rachel and Holden Diethorn, Christopher Donohue, Samantha Epstein, Stephen Farnham, Joshua Fenton, Andrei Frimu, Stephen Gorgone, Erin and Ricci Griffin, Thomas Heath, Tamir Hemo, Patrick Kanzler, Nathan Lawless, William Lippitt, Casey Necheles , Alan Robbins, Eugenia Rosu, Timothy Tribone, Erin and Bess Tripp, and Nathan Uricchio. But a special thanks to Pablo Diaz and Krystina Lynn Drasher for putting up with me through all these years. 

\vspace{-0.25\baselineskip}
Finally and above all, I would like to thank my family for their endless support, both emotional and financial. Without them, I would not have been able to accomplish any of this. 